A Toolbox for Manipulating and Assessing Color Palettes

### 2019-08-26 pkg/colorspace/paper/Toolbox/article.Rmd
incorporate improvements suggested by reviewer 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/colorspace@502 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

The \pkg{colorspace} package ships with a wide range of predefined color palettes,
specified through suitable trajectories in the HCL (Hue-Chroma-Luminance) color space.
A quick overview can be gained easily with the `hcl_palettes()` function (see Figure\ \ref{fig:hcl-palettes}):

@@

As another example for a sequential palette, we demonstrate how to create a spine plot (see the right
panel of Figure\ \ref{fig:eustockmarkets-titanic})
displaying the proportion of Titanic passengers that survived per class.
The `"Purples 3"` palette is used, which is quite similar to the \pkg{ColorBrewer.org}
[@color:Harrower+Brewer:2003] palette `"Purples"`. Here, only two colors are
employed, yielding a dark purple and a light gray.

@@

The function first transforms a given color palette to its HCL (`polarLUV()`)
coordinates. As the hues for low-chroma colors are not (or only poorly) identified,
they are smoothed by default. Also, to avoid jumps from 0 to 360 or vice
versa, the hue coordinates are shifted suitably.
By default, the resulting trajectories in the HCL spectrum are visualized by a simple line plot:

@@

>>>>

The \pkg{colorspace} package ships with a wide range of predefined color palettes,
specified through suitable trajectories in the HCL (Hue-Chroma-Luminance) color space.
A quick overview can be gained easily with the `hcl_palettes()` function (see Figure\ \ref{fig:hcl-palettes},
some of these are illustrated in more detail later):

@@

As another example for a sequential palette, we demonstrate how to create a spine plot (see the right
panel of Figure\ \ref{fig:eustockmarkets-titanic})
displaying the proportion of Titanic passengers that survived per class.
The `"Purples 3"` palette is used, which is quite similar to the \pkg{ColorBrewer.org}
[@color:Harrower+Brewer:2003] palette `"Purples"`. Here, only two colors are
employed: a dark purple that is highlighted against a light gray.

@@

The function first transforms a given color palette to its HCL (`polarLUV()`)
coordinates. As the hues for low-chroma colors are not (or only poorly) identified,
they are smoothed by default. Also, to avoid jumps from 0 to 360 or vice
versa, the hue coordinates are shifted suitably.
By default, the resulting trajectories
in the HCL spectrum are visualized by a simple line plot where the
x-axis gives the ordering of the colors in the palette. The
y-axis depicts the following information:

@@

The app has been influenced considerably by \pkg{ColorBrewer.org} [@color:Harrower+Brewer:2003].
Similarities include the selection of a qualitative, sequential, or diverging palette
from a list of predefined colors along with an example visualization. However, unlike
\pkg{ColorBrewer.org} our \pkg{shiny} app allows to tweak the HCL parameters underlying
each palette. This makes the app much more flexible but also more complex, potentially requiring more
thought and experience. Due to the flexibility, our app cannot automatically judge safety regarding
color vision deficiencies and printers/photocopiers (as \pkg{ColorBrewer.org} does) but instead it allows to emulate color vision
deficiencies and desaturation. Finally, \pkg{ColorBrewer.org} is geared towards cartography
(albeit its palettes are useful much more generally) while our \pkg{shiny} app includes a
broader range of illustrative displays.
